\documentclass{amsart}

\input{StandardPaper2.tex}

\begin{document}

\title[Pull-back Bundle Nonsense]
 {Nonsense On Pull-back Bundles and Connections}

\curraddr{}
\email{}
\date{\today}

\dedicatory{}
\subjclass[2010]{}
\keywords{}

\maketitle

\section{Vector Bundles and Sheaves of Sections}

\subsection{Vector Bundles}

Throughout, let \(\pi^E : E \to N\) be a smooth, rank \(k\), vector bundle. Give basic definitions.

\subsection{Sheaf of Sections}

\begin{defn}
Given a smooth, vector bundle \(\pi^E : E \to N\), we define the sheaf of sections \(\Gamma(E)\) as follows: for each open set \(U \subseteq N\), define the \emph{local sections} of \(E\) over \(U\) by
\[
\Gamma(U, E) = \{s : U \to E : \pi^E \circ s = \operatorname{Id}_N\}.
\]
For any pair of open sets, \(V \subseteq U \subseteq N\), define the \emph{restriction maps}, \(r_{V,U} : \Gamma(U, E) \to \Gamma(V, E)\) for \(s \in \Gamma(U, E)\) by,
\[
(r_{V,U} (s)) (x) = s(x), \quad x \in V.
\]
The sheaf of sections, \(\Gamma(E)\) is the collection \(\{\Gamma(U, E), r_{V,U}\}\) of local sections and restriction maps.
\end{defn}

\begin{rem}
The restriction maps \(r_{V,U}\) satisfy the \emph{Pre-sheaf axioms}:
\begin{enumerate}
\item \(r_{U,U} = \operatorname{Id}\).
\item For \(W \subseteq V \subseteq U\), \(r_{W,U} = r_{W,V} \circ r_{V,U}\).
\end{enumerate}
and the \emph{Sheaf axioms}:
\begin{enumerate}
\item (Locality) If \(s, t \in \Gamma(U, E)\) and \(\{U_{\alpha}\}\) is an open cover of \(U\) such that \(r_{U_{\alpha},U} s = r_{U_{\alpha},U} t\), then \(s = t\).
\item (Gluing) If \(\{U_{\alpha}\}\) is an open cover of \(U\) and there are sections \(s_{\alpha} \in \Gamma(U_{\alpha}, E)\) for each \(\alpha\) such that \(r_{U_{\alpha} \cap U_{\beta},U_{\alpha}} s_{\alpha} = r_{U_{\alpha} \cap U_{\beta},U_{\beta}} s_{\beta}\), then there exists \(s \in \Gamma(U, E)\) such that \(r_{U_{\alpha},U} s = s_{\alpha}\) for each \(\alpha\).
\end{enumerate}
\end{rem}

\begin{defn}
For vector bundles, \(E \to N\) and \(F \to N\) a \emph{local map} \(\varphi : \Gamma(E) \to \Gamma(F)\) is a collection of smooth maps, \(\varphi_U : \Gamma(E, U) \to \Gamma(F, U)\) such that,
\[
\varphi_V \circ r^E_{V, U} = r^F_{V, U} \circ \varphi_U
\]
for any pair of open sets \(V \subset U \subset N\).
\end{defn}

\begin{example}
A vector bundle map is smooth map \(\varphi : E \to F\) such that,
\[
\pi^F \circ \varphi = \pi^E
\]
and such that,
\[
\varphi_x : E_x \to F_x
\]
is linear, where \(\varphi_x = \varphi|_{E_x}\) is the restriction of \(\varphi\) to the fibre \(E_x = (\pi^E)^{-1}(x)\).

For any open set \(U\) we may then define,
\[
(\varphi_U (s)) (x) = \varphi \circ s (x) = \varphi_x(s(x))
\]
for \(x \in U\). It's a simple matter to verify that the collection \(\{\varphi_U\}\) defines a local map \(\Gamma(E) \to \Gamma(F)\).
\end{example}

\begin{rem}
The previous example shows that any vector bundle map is a local map. The converse is not true. That is, there exist local maps that are not induced by bundle maps. Note that if \(\varphi : \Gamma(E) \to \Gamma(F)\) is a local map, we might try to define,
\[
\varphi(e) = (\varphi_U s) (\pi^E(e)), \quad e \in E
\]
where \(U \subset N\) is an open neighbourhood of \(x = \pi^E(e)\) and \(s \in \Gamma(U, E)\) is any local section such that \(s(\pi^E(e)) = e\). Working in a local trivialistion, it's easy to construct such a section \(s\). The problem is that the result will in general depend on the choice of \(s\) as illustrated by the next example.
\end{rem}

\begin{example}
Let \(N = \R\) and \(E = \R \times \R\) with projection \(\pi^E = \pi^1 : \R \times \R \to \R\) onto the first factor. A local section \(s \in \Gamma(U, E)\) is a map \(s : U \to \R \times \R\) such that \(\pi^1 \circ s (x) = x\) so that we may write,
\[
s(x) = (x, f_s(x))
\]
where \(f_s : U \to \R\) is a smooth function. Define the derivative local operator \(D : \Gamma(E) \to \Gamma(E)\) by,
\[
(D s) (x) = (x, f_s'(x))
\]
where \(f_s'\) is the derivative of \(f_s\). Then it's easily checked that \(D\) is a local operator. However, value of \(D s\) at \(x\) depends on \(s\) is a full neighbourhood of \(x\) and so the attempted construction in the remark fails.

An explicit example of the failure is as follows: let \(e = (0, 0) \in E\) and define sections \(t,s \in \Gamma(\R, E)\) by
\[
s(x) = (x, x), \quad t(x) = (x, x^2).
\]
Then
\[
s(\pi^E(e)) = s(0) = (0,0) = e = t(\pi^E(e)),
\]
but
\[
(D s) (\pi^E(e)) = (0, 1) \ne (0, 0) = (D t) (\pi^E(e)).
\]
That is, we just used the simple fact that two smooth functions may have the same value at a point, but different derivatives.
\end{example}

The previous remark and example show that local maps depend on the value of a section in a neighbourhood of a point and not just the value of the section at a point. The example given is the prototypical example for us, where a local map is a differential operator. A differential operator, though not pointwise dependent, is dependent only an arbitrarily small neighbourhood of a point. We can express this notion concisely via the \emph{stalks} of a sheaf which consist of the \emph{germs} of sections at each point \(x \in N\).

\begin{defn}
A \emph{germ} of sections at \(x\) is an equivalence class \([U, s]_x\) where \(U\) is an open neighbourhood of \(x\) and \(s \in \Gamma(U, E)\) is a local section. Two pairs \((U, s)\), \((V, t)\) are equivalent if there exists an open set \(W \subset U \cap V\) with \(x \in W\) and such that
\[
r_{W, U} s = r_{W, V} t.
\]
The \emph{stalk} \(\Gamma_x(E)\) of \(E\) at \(x\) is the set of germs, \([U, s]_x\) of sections at \(x\).
\end{defn}

\begin{rem}
Recall that for \(x \in N\), the fibre \(E_x\), of \(E\) over \(x\) is,
\[
E_x = (\pi^E)^{-1} (x)
\]
which, via a local trivialisation, admits the structure of a \(k\)-dimensional vector space. Given a section \(s\), we have the fibre \(s_x = s(x) \in E_x\) of \(s\) at \(x\). Compare this with the germ - the fibre is pointwise, the germ depends on an arbitrarily small neighbourhood of \(x\).
\end{rem}

Although as noted above, local operators in general are not determined by their values at a point, it is true that they only depend on the germ at a point. To see this first let us note that local operators give rise to operators of stalks.

\begin{defn}
Let \(\varphi : \Gamma(E) \to \Gamma(F)\) be a local operator. For \(x \in N\), the \emph{stalk map} \(\varphi_x : \Gamma_x(E) \to \Gamma_x (F)\) is defined by,
\[
\varphi_x ([U, s]_x) = [U, \varphi_U(s)]_x.
\]
\end{defn}

It is an easy matter to see that the definition is well defined, independent of the choice of representative for the equivalence class. Now we may see that a local map is determined by it's stalk maps.

\begin{lemma}
Let \(\varphi, \psi : \Gamma(E) \to \Gamma(F)\) be local operators. Suppose that \(\varphi_x = \psi_x\) for every \(x \in N\). Then \(\varphi = \psi\).
\end{lemma}

\begin{proof}
Straightforward (but necessary!) exercise.
\end{proof}

\begin{rem}
If we have a vector bundle map \(\varphi : E \to F\), then we wrote \(\varphi_x\) for the fibre map and also for the stalk map. In this situation the two are in fact one and the same. This is yet another straightforward but necessary exercise.
\end{rem}

\section{Connections}

\subsection{Affine Connections}

\begin{defn}
An affine connection \(\nabla\) on a vector bundle \(\pi^E : E \to N\) is a local map
\[
\nabla : \Gamma(E) \to \Gamma(T^{\ast} N \otimes E)
\]
satisfying the Leibniz rule,
\[
\nabla (f \otimes s) = df \otimes s + f \nabla s.
\]
\end{defn}

According to the definition of a local map, for each open set \(U\) we have a map,
\[
\nabla_U : \Gamma(U, E) \to \Gamma(U, T^{\ast} N \otimes E)
\]
compatible with the restriction maps so that,
\[
r_{V, U} \nabla_U s = \nabla_V r_{V, U} s
\]
whenever \(V \subseteq U\) and \(s \in \Gamma(U, E)\). The Leibniz rule here means that for any \(s \in \Gamma(U, E)\) and smooth function \(f : U \to \R\) we have,
\[
\nabla_U (fs) = d_U f \otimes s + f \nabla_U s
\]
where we recall that the differential \(d\) is a local map \(\Gamma(C^{\infty}(N)) \to \Gamma(T^{\ast} N)\), and the tensor product \(C^{\infty}(N, \R) \otimes E \simeq E\) with \(f \otimes s\) denoting the section \(x \mapsto f(x) s(x) \in E\).

\begin{rem}
Since \(\nabla s \in \Gamma(T^{\ast} N \otimes E)\), we can contract the \(T^{\ast} N\) part of \(\nabla s\) with a tangent vector \(X \in TN\), the result of which is in \(\Gamma(E)\). We write this contraction various ways:
\[
\nabla s (X) = \nabla_X s = \Tr (\nabla s \otimes X).
\]

By definition, \(\nabla_X s\) depends tensorialy on \(X\) and hence depends pointwise on \(X\) so that for example, if \(X, Y \in \Gamma(TM)\) such that \(X(x) = Y(x)\) then,
\[
(\nabla_X s) (x) = (\nabla_Y s) (x).
\]
Moreover, \((\nabla_X s) (x)\) depends only the stalk, \([s]_x\) of \(s\) at \(x\). Since \(\nabla\) is local however, and (because of the Leibniz rule) not a bundle map, \(\nabla_X s\) cannot be determined from the fibre, \(s_x\) of \(s\) at \(x\).
\end{rem}

\subsection{Ehresmann Connections}

There is way of interpreting connections a little differently, in terms of \emph{Ehresmann Connections}. For a smooth, vector bundle \(\pi^E : E \to N\), the map \(\pi^E\) is, in particular a smooth map hence has a differential, \(d\pi^E : TE \to TN\). As a smooth manifold \(E\) has dimension \(k+n\) where \(n = \operatorname{dim} (N)\). As a smooth map, \(\pi^E\) is a submersion hence \(d\pi^E\) has constant rank \(k\) everywhere and it's kernel is a smooth vector sub-bundle of \(TE\) over \(E\).

\begin{defn}
The \emph{vertical bundle} \(VE\) of \(TE\) is defined as
\[
VE = \operatorname{ker} d\pi^E.
\]
A \emph{horizontal bundle} is any complmentary sub-bundle \(HE \subseteq TE\) to \(VE\); that is,
\[
TE \simeq VE \oplus HE.
\]
\end{defn}

\begin{defn}
A \emph{Ehresmann Connection} \(H\) on \(E\) is a horizontal bundle of \(TE\).
\end{defn}

The next lemma relates the notion of Affine Connection \(D\) to horizontal sub-bundle \(H\).

\begin{lemma}
There is a one-to-one correspondence between Affine Connections \(D\) on \(E\) and horizontal sub-bundles \(H\) of \(TE\).
\end{lemma}

\begin{proof}

\end{proof}

\subsection{Connections as Sections of Jet Bundles}

We need yet more theory, interpreting connections as sections of \emph{Jet Bundles}.

\begin{defn}
Jet Bundle.
\end{defn}

\section{Curvature}

For affine connections:

\begin{defn}
\[
R(X, Y) s = D_X D_Y s - D_Y D_X s - D_{[X, Y]} s.
\]
\end{defn}

For Ehrassmen connections:

\begin{defn}
Maurer-Cartan Form.
\end{defn}

For Jet Bundles:

\begin{defn}
Not sure yet.
\end{defn}

\section{Pull Back Bundle}

\begin{lemma}[Pull-back Bundle]
There exists a unique (up to vector bundle isomorphism) smooth vector bundle, the \emph{pull-back bundle} \(f^{\ast} \pi : f^{\ast}E \to M\) whose sections over an open set \(U  \subseteq M\) are
\[
\Gamma(U, f^{\ast} E) = \{s : U \to E | \pi^E \circ s = f\}.
\]
\end{lemma}

\begin{defn}
Given an open set \(V \subseteq N\) and a section \(s \in \Gamma(V, E)\) of \(E\), we define the pull-back \(f^{\ast} s \in \Gamma(f^{-1}(V), f^{\ast} E)\) by
\[
f^{\ast} (x) = s(f(x)), \quad x \in f^{-1}(V).
\]
\end{defn}

\begin{rem}
In general, a local section \(S\) of \(f^{\ast} E\) is not equal to any pull-back section \(f^{\ast} s\). It is true however, that the sections of \(f^{\ast} E\) are \emph{locally generated} by pull-back sections: any open set \(U \subseteq M\) is covered by sets \(U_{\alpha} = f^{-1}(V_{\alpha}) \cap U\) where \(V_{\alpha}\) trivialise \(E\). Then for each \(\alpha\), we have a local frame \(\{s^{\alpha}_i\}_{i=1}^k\) for \(E\) over \(V_{\alpha}\) and any section \(S \in \Gamma(U, f^{\ast} E)\) may be written over \(U_{\alpha}\) as
\[
S(x) = G^i(x) f^{\ast} s^{\alpha}_i(x)
\]
for smooth functions \(G^i \in C^{\infty}(U_{\alpha}, \R)\). It's important to note here that in general, \(G^i \ne f^{\ast} g^i\) for smooth functions \(g^i\) - otherwise we would have \(S = f^{\ast} s\) where \(s = g^i s^{\alpha}_i\).

In the particular case that \(f\) is an immersion, any smooth function \(G : U \to \R\) may be locally extended to \(g : V \to \R\) in the sense that \(G = f^{\ast} g\) along the open set \(f^{-1}(V) \cap U\), but not in general globally. If \(f\) is an embedding, then the extension \(G = f^{\ast} g\) for \(g\) a smooth function \(N \to \R\) can be done globally. Thus, in the case of an immersion, all sections \(S \in \Gamma(U, f^{\ast} E)\) are locally pull-backs \(S = f^{\ast} s\). For an embedding, all sections \(S \in \Gamma(U, f^{\ast} E)\) are globally pull-backs \(S = f^{\ast} s\).

Once more, let us emphasise the point that for a general smooth map, \(f: M \to N\) sections of the pull-back bundle cannot even be written locally as pull-back sections but as a linear combination of pull back sections where the coefficients are smooth functions on \(M\).
\end{rem}

\begin{example}
A particular example is the tangent bundle, \(TN \to N\). In this case, we get a bundle map,
\begin{align*}
f_{\ast} : TM &\to f^{\ast} TN \\
X &\mapsto df \cdot X,
\end{align*}
that is,
\[
(f_{\ast} X) (x) = df_{x} \cdot X(x), \quad x \in M.
\]

As noted above, \(f^{\ast} X\) so defined will not in general be extendable, even locally to a vector field on \(N\), but it does define a section of the pull-back bundle. If \(f\) is an embedding, \(f_{\ast} X\) may be extended globally to vector field on \(N\), and if \(f\) is an immersion, such an extension is possible locally.

For immersions, by definition \(df_x\) is injective at each \(x\) and so we see that \(f_{\ast}\) is an injective as a bundle map and hence,
\[
f_{\ast} : TM \hookrightarrow f^{\ast} TN
\]
realises \(TM\) as isomorphic to the sub-bundle of \(f_{\ast} (TM) \subseteq f^{\ast} TN\).

The normal bundle \(\nu M\) is defined as the quotient bundle \(f^{\ast} TN/TM\) through the exact sequence,
\[
0 \to TM \overset{f_{\ast}}{\to} f^{\ast} TN \to \nu M \to 0
\]
of vector bundles

If \(TN\) has a metric \(g\), then the metric splits the exact sequence. That is we have the orthogonal decomposition,
\[
f^{\ast} TN \simeq TM \oplus \nu M
\]
and an inclusion,
\[
\nu M \hookrightarrow f^{\ast} TN
\]
with
\[
\nu M = \{Z \in f^{\ast}TN : g(Z, f_{\ast} X) = 0 \quad \forall X \in TM\}.
\]
\end{example}

\section{Connections on Pull-back Bundles}

Suppose now our vector bundle \(\pi^E : E \to N\) has a connection \(D\). There are several ways to define the pull back connection \(f^{\ast} D\) on \(f^{\ast} E\), but whatever method we choose, the pull-back bundle may be characterised by a defining property.

\begin{lemma}
\label{lem:pullback_connection}
There is a unique connection, \(f^{\ast} D\), the \emph{pull-back connection} on the vector bundle \(\pi^{f^{\ast} E} : f^{\ast}E \to M\) satisfying
\[
(f^{\ast} D)_X f^{\ast} s = D_{f_{\ast} X} s
\]
for any tangent vector \(X \in TM\) an any local section \(s \in \Gamma(V, E)\).
\end{lemma}

The point now is to show,
\begin{enumerate}
\item \(f^{\ast} D\) is well defined,
\item \(f^{\ast} D\) is the unique connection with the above property.
\end{enumerate}

First, let us prove uniqueness.

\begin{proof}[Proof of Uniqueness]
Suppose we have a connection \(\nabla\) on \(f^{\ast} E\) that acts on pull-back sections by,
\[
\nabla_X f^{\ast} s = D_{f_{\ast} X} s.
\]
For an arbitrary section \(S \in f^{\ast} E\), choose a local frame \(\{s_i\} \in \Gamma(V, E)\) and write,
\[
S = G^i f^{\ast} s_i.
\]
By the Leibniz rule, and the action of \(\nabla\) on pull-back sections, we must have
\[
\nabla_X S = X (G^i) f^{\ast} s_i + G_i D_{f_{\ast} X} s_i.
\]
The right hand side uniquely determines \(\nabla\).
\end{proof}

We offer up three proofs for existence.

\begin{proof}[Proof of \Cref{lem:pullback_connection} by local generation]
As in the proof of uniqueness, for an arbitrary section \(S \in f^{\ast} E\), choose a local frame \(\{s_i\} \in \Gamma(V, E)\) and write,
\[
S = G^i f^{\ast} s_i.
\]
and define,
\[
(f^{\ast} D)_X S = X (G^i) f^{\ast} s_i + G_i D_{f_{\ast} X} s_i.
\]

We need to verify \(f^{\ast} D\) is well defined. That is, if we choose another frame \(\{s'_i\}\) and write, \(S = (G')^i f^{\ast} s\_i\) then we need
\[
X (G^i) f^{\ast} s_i + G^i D_{f_{\ast} X} s_i = X ((G')^i) f^{\ast} s'_i + (G')^i D_{f_{\ast} X} s'_i.
\]
Our two frames \(\{s_i\}\) and \(\{s'_i\}\) are related by a change of frame, \(s'_i = \tau^j_i s_j\) and the result follows since \(D\) is a connection ensuring the right hand side transfoms correctly to the left hand side. This is one of those situations where the reader should check this last claim for themselves to truly believe it!
\end{proof}

The next proof is in terms of Ehresmann Connections.

\begin{proof}[Proof of \Cref{lem:pullback_connection} by horizontal sub-bundles]
For an affine conection \(D\), let \(H_D\) denote the corresponding horizontal sub-bundle. Then \(f^{\ast} D\) is the affine connection corresponding to the horizontal sub-bundle, \(f^{\ast} H\).
\end{proof}

Finally, our third proof is in terms of Jet Bundles.

\begin{proof}[Proof of \Cref{lem:pullback_connection} by jets]
For an affine connection, \(D\), let \(J_D\) be the corresponding section of \(J_1(E) \to N\). The pull-back connection is simply the affine connection determined by the pull-back section,
\[
f^{\ast} J_D \in \Gamma(f^{\ast} J_1(E)) = \Gamma(J_1(f^{\ast}(E))).
\]
\end{proof}

\section{Extrinsic Curvature}
\end{document}
